---
title: 'PML: Course Project'
author: "Yuri Isakov"
date: "28 January 2016"
output: html_document
---

### Obtaining data

Training and test set was downloaded into working directory. Then empty columns were excluded. Actually it could be done in more elegant way during loading csv files.
```{r}
train <- read.csv("pml-training.csv", stringsAsFactors = FALSE)
test <- read.csv("pml-testing.csv", stringsAsFactors = FALSE)
minusCol1 <- colSums(train == "") > 10000
minusCol1[is.na(minusCol1)]  <- TRUE
minusCol2 <- colSums(is.na(train)) > 10000
minusCol <- minusCol1 | minusCol2
minusCol[c(1, 3:6)] <- TRUE
train <- train[,!minusCol]
```

#### Using `user_name` variable
I saw on forums that it's not necessary to use ID variable. I believe different persons can do things in different ways. That's why I didn't omit this variable. Here's a plot that shows my motivation. For some users parameters may be significant different.

```{r cache=TRUE, message=FALSE}
qplot(train[,1], train[,2], colour = train$classe)
```


It could be the shortest report ever because all is need to run this code:
```{r cache=TRUE, message=FALSE}
library(caret)
# fit_gbm <- train(classe ~., data = tr, method = "gbm")
fit_gbm <- readRDS("gbm01.rds")
confusionMatrix(predict(fit_gbm, train), train$classe)
```

It gives 20/20 answers. While we have correct answers and some time let's try something else just for fun.

